Cache is King: Presenting a Suite of Fast, Correlation Function Codes

### 2015-03-27 paper/ms.tex
Updated the logic in DDrppi to remove using the comoving distance array. Now, I check for max. separation before proceeding with the division

%%% Definitions added by Manodeep 
\newcommand{\Msun}{M_{\odot}}
\newcommand{\hMsun}{h^{-1}M_{\odot}}
\newcommand{\sdss}{{\small{SDSS}}\xspace}
\newcommand{\hMpc}{\ensuremath{{h^{-1}Mpc}\xspace}}
\newcommand{\hkpc}{\ensuremath{{h^{-1}kpc}\xspace}}
\newcommand{\emcee}{\texttt emcee}
\newcommand{\rmax}{\ensuremath{{r_{max}}}\xspace}
\newcommand{\xir}{\ensuremath{{DD(r)}}\xspace}
\newcommand{\xiofr}{\ensuremath{{\xi(r)}}\xspace}
\newcommand{\wprp}{\ensuremath{{w_p(r_p)}}\xspace}
\newcommand{\xirppi}{\ensuremath{{\xi(r_p,\pi)}}\xspace}
\newcommand{\todo}[1]{\marginpar{TODO}{\color{red}#1}}

@@

\begin{abstract}
We have entered the era of `Big Data' where we are gathering a tremendous amount of data that
needs to be processed as well as modeled accurately. In particular, large galaxy surveys
like the Sloan Digitial Sky Survey requires computing correlation functions. While measuring the
correlation function in the data is not a bottle-neck, modeling the observed galaxy distribution correctly
{\em} requires a MCMC chain and repeated measurements of $\xi(r)$ and/or $\xi(r_p,\pi)$. However, 
the architecture of modern CPU's means that best performance is only obtained when cache misses 
are kept to a minimum. Here, I present a suite of 3 correlation function codes that exploit 
current CPU architecture and hand-written AVX

The accompanying codes are publicly available at \url{https://bitbucket.org/manodeep/Corrfunc/}. 
\end{abstract}

@@

Modern cpu's have a hierarchy of memory locations; the smallest and fastest are closest to the cores while the largest and slowest are the
farthest. All cpu instructions need to be carried out from cpu registers - there are $\sim 10-30$ registers typically available and the 
access times can be thought of as instantaneous. Next up is the $L1$ cache divided into $L1D$ for data and $L1I$ for instructions cache. Since 
the cpu always necessarily executes instructions that are close together, we will ignore the instruction cache from now on. Typical $L1$ cache 
sizes range from 64KB to 128 KB. Next level up is the $L2$ cache, typically $\sim$ 256KB to 1 MB. The last level cache or the $L3$ cache 
is usually shared across all cores on the socket and can be 10 MB to 40 MB.

@@

\subsection{\xir}
\subsection{\xirppi}
\subsection{\wprp}

>>>>

%%% Definitions added by Manodeep 
\newcommand{\Msun}{M_{\odot}}
\newcommand{\hMsun}{h^{-1}M_{\odot}}
\newcommand{\sdss}{{\small{SDSS}}\xspace}
\newcommand{\hMpc}{\ensuremath{{h^{-1}Mpc}\xspace}}
\newcommand{\hkpc}{\ensuremath{{h^{-1}kpc}\xspace}}
\newcommand{\emcee}{\texttt emcee}
\newcommand{\rmax}{\ensuremath{{r_{max}}}\xspace}
\newcommand{\xir}{\ensuremath{{DD(r)}}\xspace}
\newcommand{\xiofr}{\ensuremath{{\xi(r)}}\xspace}
\newcommand{\wprp}{\ensuremath{{w_p(r_p)}}\xspace}
\newcommand{\xirppi}{\ensuremath{{\xi(r_p,\pi)}}\xspace}
\newcommand{\cis}{\ensuremath{{pN(r)}}\xspace}
\newcommand{\ddrppi}{\ensuremath{{DD(r_p,\pi)}}\xspace}
%% \newcommand{\ddrppi}{\ensuremath{{DD(r_p,\pi)}}\xspace}
\newcommand{\wtheta}{\ensuremath{{DD(\theta)}}\xspace}
\newcommand{\todo}[1]{\marginpar{TODO}{\color{red}#1}}

@@

\begin{abstract}
In the era of `Big Data', we need special computing techniques to both process the raw data as well 
generate and evaluate models to test our theories. In particular, large galaxy surveys
like the Sloan Digitial Sky Survey requires computing a variety of $n$-point statistics, e.g., the 
2-pt projected correlation function, $\wprp$. While measuring the correlation function in the data is not a bottle-neck, 
modeling the observed galaxy distribution correctly
{\em requires} a MCMC chain and repeated measurements of $\wprp$. To efficiently 
measure the correlation functions, we need to write code that accounts for cache hierarchies while 
simultaneously exploits the wide vector registers present in the relatively modern CPUs.  
Here, I present a suite of OpenMP parallelized, clustering codes that exploit current CPU micro-architecture with hand-written AVX 
intrinsics. Given a 3-D distribution of points in a cartesian system, these codes compute spatial clustering statistics 3-D correlation functions \xir, \xiofr, \xirppi; 2-point 
projected correlation function \wprp; counts-in-spheres \cis. For a 3-D particle distribution on the sky, the codes can compute 
projected correlation function \ddrppi, angular correlation function \wtheta as well as \cis. These codes are designed to be blazing fast 
and can compute \wprp for $\mathcal{O}$(1 million) galaxies in $\sim 6$ seconds on a post-2011 CPU, which is a factor of few faster than 
the existing public correlation function routines.  
The accompanying codes are publicly available at \url{https://bitbucket.org/manodeep/Corrfunc/}. 
\end{abstract}

@@

Modern cpu's have a hierarchy of memory locations; the smallest (and fastest) are physically located close to the computing cores 
while the largest (and slowest) are the farthest. All cpu instructions need to be carried out from cpu registers - there are $\sim 100$ 
registers typically available and the 
access times can be thought of as instantaneous. Next up is the $L1$ cache divided into $L1D$ for data and $L1I$ for instructions cache. Since 
the cpu always necessarily executes instructions that are close together, we will ignore the instruction cache from now on. Typical $L1$ cache 
sizes range from 64KB to 128 KB (shared between instruction and data). 
Next level up is the $L2$ cache, typically $\sim$ 256KB to 1 MB. The last level cache or the $L3$ cache is usually shared across all 
cores on the socket and can be 10 MB to 40 MB.

@@

\subsection{\xir}
\subsection{\xirppi}
\subsection{\wprp}
\subsection{}
